Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2016-07-23 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
editting
Fixed refs for primer and paint.  other edits

After receipt of the gypsum wallboard panels, the panels were laid out on the floor of the NIST Large Fire Research Laboratory, finished paper side up and painted with a roller.  The first coat was a latex primer, http://buyat.ppg.com/rep\_pafpainttools\_files/Pghpaints/tdb/78200.pdf.  Once the primer dried a top coat of white latex paint was applied http://buyat.ppg.com/rep\_pafpainttools\_files/Olympic/tdb/78010.pdf.  After the panels dried, they were weighed. The average mass of paint on the finished surface of the gypsum wallboard was 0.33~kg. Then the paint thickness was estimated an ultrasonic coating thickness gauge~\cite{defelsko}.  Over the course of the experiments, random painted panels where checked with the ultra sonic device. The thickness of the paint varied from 0.05~mm to 0.13~mm.  Some of the painted panels were then cut-up into samples for the bench-scale experiments.

@@

@@

The fire pattern experiments were conducted in residential scale compartments built for the purpose of these experiments.  The compartments were constructed on top of a portable wood floor platform.  Each of the test compartments consisted of three 3.6~m long by 2.4~m high wood framed walls. The wood elements of the frame were composed of kiln-dried fir with a nominal cross section of 90~mm by 38~mm.  A partial ceiling with a width of at least 1.2~m, measured from the back wall of the compartment was installed across the width of the compartment, this allowed for the plume to impact the ceiling and form a ceiling jet along the back wall, before the fire gases flowed up and out of the compartment and into the 6~m by 6~m hood in the NIST Large Fire Research Laboratory. The interior surface of the walls and ceiling was composed of 12.7~mm thick, regular core gypsum wallboard. As defined by ASTM, gypsum wallboard is designed for use on walls, ceilings, or partitions.  It is composed of a non-combustible core, essentially gypsum, with paper bonded to the surface of the core. The back wall of the compartment consisted of three full gypsum board panels, each 1.2~m wide and 2.4~m high.  The front wall of the compartment was open along the full width of the compartment to minimize the impact of entrained air being channeled through a doorway. These experiments were conducted without instrumentation in the walls for several reasons; to compare the non-instrumented gypsum wallboard fire pattern with the instrumented gypsum wallboard fire pattern results to determine if the instrumentation affected the pattern generation, to provide more realistic conditions, and to enable the conduct of additional replicate experiments given the limited laboratory availability. A photograph of the compartment fire experimental arrangement is shown in Figure~\ref{Compartment_Test}. In the instrumented gypsum wallboard experiments, as the burners were moved away from the position against the wall, the fire effects on the wall diminished significantly.  Therefore most of the fire pattern compartment experiments were conducted with the burners against the wall or corner of the compartment or less than one burner length from the wall.

@@

Values generated in the bench scale experiments will be utilized as input to predict real scale results using the algorithms.  Then the predicted results will be compared with the full-scale experiment results.

>>>>

After receipt of the gypsum wallboard panels, the panels were laid out on the floor of the NIST Large Fire Research Laboratory, finished paper side up and painted with a roller.  The first coat was a latex primer~\cite{Latex_Primer}.  Once the primer dried a top coat of white latex paint was applied ~\cite{Latex_Paint}.  After the panels dried, they were weighed. The average mass of paint on the finished surface of the gypsum wallboard was 0.33~kg. Then the paint thickness was estimated an ultrasonic coating thickness gauge~\cite{defelsko}.  Over the course of the experiments, random painted panels where checked with the ultra sonic device. The thickness of the paint varied from 0.05~mm to 0.13~mm.  Some of the painted panels were then cut-up into samples for the bench-scale experiments.

@@

The purpose of this chapter is to develop and compare data from bench scale apparatus that could be used to analyze fire damage to gypsum wallboard.
This type of testing might be practical for a fire investigator to use based on the low cost, relative to a full-scale fire test, and the small size of the samples needed for the tests.  In the previous sections of this study, burners were used to examine the repeatibility of each fire's characteristics, the amount of heat transfered to a wall as a function of distance and the damage to a section of painted gypsum wallboard.  While the data generated from those experiments provide valuable insight, it would be of valauble to be able to develop assessments of fire patterns based on data from one or a series of bench scale tests.

@@

The fire pattern experiments were conducted in residential scale compartments built for the purpose of these experiments.  The compartments were constructed on top of a portable wood floor platform.  Each of the test compartments consisted of three 3.6~m long by 2.4~m high wood framed walls. The wood elements of the frame were composed of kiln-dried fir with a nominal cross section of 90~mm by 38~mm.  A partial ceiling with a width of at least 1.2~m, measured from the back wall of the compartment was installed across the width of the compartment, this allowed for the plume to impact the ceiling and form a ceiling jet along the back wall, before the fire gases flowed up and out of the compartment and into the 6~m by 6~m hood in the NIST Large Fire Research Laboratory. The interior surface of the walls and ceiling was composed of 12.7~mm thick, regular core gypsum wallboard. As defined by ASTM, gypsum wallboard is designed for use on walls, ceilings, or partitions.  It is composed of a non-combustible core, essentially gypsum, with paper bonded to the surface of the core. The back wall of the compartment consisted of three full gypsum board panels, each 1.2~m wide and 2.4~m high.  The front wall of the compartment was open along the full width of the compartment to minimize the impact of entrained air being channeled through a doorway. These experiments were conducted without instrumentation in the walls for several reasons:
\begin{enumerate}
\item to compare the non-instrumented gypsum wallboard fire pattern with the instrumented gypsum wallboard fire pattern results to determine if the instrumentation affected the pattern generation
\item to provide more realistic conditions
\item to enable the conduct of additional replicate experiments
\end{enumerate}

A photograph of the compartment fire experimental arrangement is shown in Figure~\ref{Compartment_Test}. In the instrumented gypsum wallboard experiments, as the burners were moved away from the position against the wall, the fire effects on the wall diminished significantly.  Therefore most of the fire pattern compartment experiments were conducted with the burners against the wall or corner of the compartment or less than one burner length from the wall.

@@

Values generated in the bench scale experiments will be utilized as input to predict real scale results using the algorithms.  Then the predicted results will be compared with the full-scale experimental results.
